---
title: "ARIMAX Modeling: Ice Cream Consumption"
author: "Andira Putri"
output: pdf_document
---

In this project, we do ten exercises from: https://www.r-bloggers.com/forecasting-arimax-model-exercises-part-5/. We will be working on the `Icecream` data set from the `Ecdat` package.

```{r,message=FALSE}
#install.packages("Ecdat")
library(Ecdat)
data(Icecream)
```

#### Exercise 1. Load the dataset, and plot the variables `cons` (ice cream consumption), `temp` (temperature), and `income`.

```{r}
t=1:30 #time values
df=cbind(t,Icecream) #add time values to dataset
library(GGally)
ggpairs(df)
```

#### Exercise 2. Estimate an ARIMA model for the data on ice cream consumption using the auto.arima function. Then pass the model as input to the forecast function to get a forecast for the next 6 periods (both functions are from the forecast package).

#### Exercise 3.

#### Exercise 4.

#### Exercise 5.

#### Exercise 6.

#### Exercise 7.

#### Exercise 8.

#### Exercise 9.

#### Exercise 10.


