---
title: "ARIMAX Modeling - US Economy"
author: "Andira Putri"
output: pdf_document
---

This exercise is taken from **Time Series Analysis and Its Applications: With R Examples** by Shumway and Stoffer. We will be using the `econ5` data set from the `astsa` library. `econ5` is a five quarterly economic series containing the following numeric variables: quarterly U.S. unemployment, GNP, consumption, government investment, and private investment. There are 161 observtions spanning from 1948-III to 1988-II.

Consider the data set `econ5`. The seasonal component has been removed from the data. Concentrating on unemployment $(U_t)$, GNP $(G_t)$, and consumption $(C_t)$, fit a vector ARMA model to the data after first logging each series, and then removing the linear trend. That is, fit a vector ARMA model to $x_t=(x_{1t},x_{2t},x_{3t})^t$ where, for example, $x_{1t}=log(U_t)-\hat{\beta_0}-\hat{\beta_1}t$, where $\hat{\beta_0}$ and $\hat{\beta_1}$ are the least squares estimates for the regression of $log(U_t)$ on time t. Run a complete set of diagnostics on the residuals.

```{r,include=FALSE}
#for data set
library(astsa)
data(econ5)
#for plots
library(gridExtra)
library(ggplot2)
#for auto.arima and checkresiduals
library(forecast)
```

#### Curious Plotting

```{r,echo=FALSE}
# create data frame for variables of interest
t=c(1:161)
u=econ5$unemp
g=econ5$gnp
c=econ5$consum
df=data.frame(t,u,g,c)
# exploratory plots
p1<-ggplot(df,aes(t,u))+geom_jitter()+ggtitle("Unemployment vs. Time")
p2<-ggplot(df,aes(t,g))+geom_jitter()+ggtitle("GNP vs. Time")
p3<-ggplot(df,aes(t,c))+geom_jitter()+ggtitle("Consumption vs. Time")
grid.arrange(p1,p2,p3,ncol=2)
```

From the plots, GNP and consumption would be interesting variables to use as regressors. We can use time, GNP, and consumption to predict unemployment in the United States.

#### Model Fitting

```{r}
# log transform
log.u=log(u)
log.g=log(g)
log.c=log(c)
log.df=cbind(t,log.u,log.g,log.c)
# fit vector ARMA model
varma=auto.arima(log.u,xreg=log.g+log.c)
summary(varma)
```

#### Diagnostics

```{r,echo=FALSE}
checkresiduals(varma)
```

Our residuals are normally-distributed for the most part, and there is no clear spike in the ACF diagram. The Ljung-Box test has a p-value of 0.05297 at the 5% level, so the data is independently distributed and there are no remarkable autocorrelations at any lag. This is a desirable result.

#### Forecasting

We randomly generate vectors for `log(g)` and `log(c)` to use as input values for our forecasting model. Both vectors are of length 8, so we aim to forecast the next 8 quarters' `log(u)` values. Note that for multiple regressors, we have to combine the vectors into a matrix for our `forecast()` function to work properly.

```{r,echo=FALSE}
# create matrix of regressor values
log.g_fc<-c(8.3,8.31,8.32,8.33,8.34,8.35,8.36,8.366)
log.c_fc<-c(7.86,7.862,7.869,7.9,7.92,7.925,8,8.16)
fc=forecast(varma,xreg=log.g_fc+log.c_fc)
autoplot(fc)
```